Viscous Transport in Eroding Porous Media

### 2019-08-26 docs/transport/erosion_transport.tex
worked on second numerical example

The first two of these metrics can be defined in terms of statistical
moments of trajectories.  Depending on the application, transport at the
microscale can be modelled as pure advection~\cite{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019}, and advection
dispersion equation~\cite{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\cite{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume that the trajectories are governed purely by
advection (ie.~no diffusion).  Therefore, a trajectory $\ss(t)$
initialized at $\ss_0$ is governed by
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
  \label{eqn:tracers}
\end{align}
where $\uu$ is the fluid velocity.  Since there is no diffusion, all
dispersion is a result of the particle spreading induced by the porous
geometry.  Since there is no diffusion, we apply a high-order explicit
Runge-Kutta time stepping method to compute tracer trajectories.
However, other groups have simulated trajectories using continuous time
random walks~\cite{mon-wei1965, den-cor-sch-ber2004, leb-den-car2008,
ber-cor-den-sch2006} and a Fokker-Planck equation~\cite{ica-den2019}.

@@

The tortuosity is a first statistical moment of the tracer trajectories.
However, to measure spread of trajectories, second-order moments are
computed. By computing the variance of the lengths of trajectories in a
porous media, non-local anomalous dispersion is
observed~\cite{kan-dea-nun-bij-blu-jua2014, cus-hu-den1995,
dea-leb-den-tar-bol-dav2013}, resulting in super-dispersion.  The
dispersion depends on the amount of time trajectories spend in low and
high velocity regimes~\cite{ber-sch2001}.  Anomalous dispersion has been
investigated for a variety of porous geometries, but not in the context
of eroding porous geometries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{sec:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  In our previous
work~\cite{qua-moo2018}, we examined the effect of erosion on
the area fraction, flow rate, and the total drag.  However, to
characerize transport, other quantities must be examined.  Here, we
compute the anomalous dispersion rate, the tortuosity, and the
distribution of the gap sizes.  The first two metrics can be defined in
terms of tracer trajectories governed by an advection
equation~\eqref{eqn:tracers}.

@@

We calculate the anomalous dispersion rates by analyzing the tracer
trajectories governed by equation~\eqref{eqn:tracers} for geometries
that are formed by erosion.  Given a set of trajectories, we define
$\lambda_j(t)$ to be the arclength of the trajectory
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p.
\end{align}
Then, the first and second ensemble moments are
\begin{align}
  \label{eqn:moments}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
Since we compute second-order statistics, we use 1000 tracers so that
the statistics have converged~\cite{bel-sal-rin1992}.  Then, particle
dispersion is characterized as $\sigma_\lambda$.  At early times, the
particles have not explored much of the geometry, and we expect a
ballistic motion $\sigma_\lambda \sim t$.  However, as the particles
pass the grains, their trajectories are altered, and we expect that
$\sigma_\lambda \sim t^\alpha$, with $\alpha \in (0.5,1)$, indicating
that the flow is super-dispersive.

@@

To compute the tortuosity and the anomalous dispersion rates
(Section~\ref{sec:transport}), we require trajectories of passive
tracers, governed by equation~\eqref{eqn:tracers}.  To reliably
characterize the transport, it is crucial that the trajectories are
computed accurately.  If a low-order time stepping method is used, or if
$\uu(\ss(t))$ is inaccurate, then a trajectory $\ss(t)$ can unphysically
enter a grain, rendering the trajectory meaningless to quantify
transport.  However, if we simply ignored trajectories that pass close
to grains, this would considerably skew the characterization of the
transport. Therefore, we use high-order quadrature and time stepping
methods.  In particular, depending on the proximity of $\ss(t)$ to
$\bd\Omega$ (Section~\ref{sec:fmm}, we apply the trapezoid rule or the
Barycentric quadrature rule.  For time stepping, we use a fourth-order
explicit Runge-Kutta method.  Since the trajectories are computed with
high-order methods in both space and time,  we are able to simulate
dynamics very close to the eroding bodies (see
Figures~\ref{fig:Eroding20tracer} and~\ref{fig:Eroding100tort}).

@@

The common characteristic of each of the experiments is near-contact
between the eroding bodies, outer walls, and passive tracers. We use our
numerical methods to describe and visualize the following examples:
\begin{itemize}
  \item{\bf Single Body Close to a Wall}: We consider a single eroding
  body close to the outer wall at $y=-1$.  Using a shear flow that
  vanishes on $y=-1$, we compare the eroding body's shape to a similar
  experiment of Spagnolie and Mitchell~\cite{mit-spa2017}.

  \item{\bf 20 Bodies at a Medium Porosity}: We consider 20 eroding
  bodies with a medium initial porosity.  After computing accurate
  tracer trajectories, the tortuosity and anomalous dispersion rates are
  computed and compared to those of an open channel.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points and a time step size of $\Delta t = 10^{-4}$.  The background
flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 2.  We start by analyzing the effect of
erosion on the area fraction and the flow rate.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  Initially, 62\% of the geometry is eroded
bodies corresponding to a porosity of $0.38$, and the area fraction
decreases to 0 as the bodies erode.  The trend of the area fraction
resembles that of our previous work~\cite{qua-moo2018} (Figure 10(a)),
but with a larger initial area fraction.  In
Figure~\ref{fig:Eroding20flowrate}(b), we plot flow rate $U$ required to
maintain a constant pressure drop across the channel.  Again, the trend
of $U$ resembles that of our previous work~\cite{qua-moo2018} (Figure
10(b)), except that the initial flow rate is an order of magnitude
smaller because of the larger initial area fraction.  Starting around
normalized time $0.2$, Figure~\ref{fig:Eroding20flowrate}(b) is roughly
linear which indicates that the flow rate can be written as a power law.
Using a line of best fit, the flow rate is approximately 
\begin{align} 
  U \approx \left(\frac{t}{t_f}\right)^{3.43},
\end{align}
which is the dashed line in Figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}[H]
\begin{subfigure}[b]{0.55\textwidth}
\includegraphics*[width =\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/flow_rate20dense}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a): The area fraction of 20
nearly touching eroding bodies versus normalized time. (b): The flow
rate, $U$, for a fixed pressure drop across the channel versus
normalized time.  The flow rate is initially very small, but it
eventually increases as a power law (dashed line) towards the flow rate
$U=1$ that occurs once all the bodies have eroded.}
\end{figure}

Snapshots of the bodies at four equispaced times are shown in
Figure~\ref{fig:Eroding20vort}.  The color is the vorticity which is
equivalent to the shear stress when restricted to the eroding bodies.
Therefore, erosion is fastest in regions where the magnitude of the
vorticity is largest.  Initially, several of the eroding bodies are
closer to the outer wall than the $5h$ threshold required to guarantee a
machine precision quadrature error for the trapezoid rule.  In
particular, with $h=L_2/N_\out$, the distance between bodies 1, 6, 13,
and 15 and the outer wall are $1.3h$, $2.9h$, $2.8h$, and $1.3h$,
respectively.  In addition, the distance between several pairs of
eroding bodies is too small to be accurately resolved with the trapezoid
rule at the prescribed resolution.  Particular pairs include bodies 1 \&
6, 3 \& 9, 6 \& 8, and 14 \& 18.  By using the Barycentric quadrature
rule, the interaction between these nearly-touching bodies can be
accurately resolved at the prescribed resolution.

@@

In~\cite{qua-moo2018}, we observed that erosion causes the space between
nearly touching bodies to quickly expand, and flat faces develop along
the region of near contact.  This qualitative behavior is present in
Figure~\ref{fig:Eroding20vort} between bodies 3 \& 4, 15 \& 16, and
others.  However, now that we can resolve bodies that are much closer
together, we observe that, at least initially, very little erosion
occurs between certain pairs of bodies.  For instance the opening
between bodies 1 \& 6, 3 \& 9, and 5 \& 13 grow much slower than the
opening between bodies 15 \& 16.  A common feature of the openings that
grow slowly is that they are nearly perpendicular to the main flow
direction, and this results in very low erosion because of the slow flow
between the bodies.

We also observe that erosion creates a network of channels from the
inlet to outlet where the velocity and vorticity, and therefore erosion
rate, are much larger when compared to other regions.  These channels
can be further visualized by considering the streamlines.  In
Figure~\ref{fig:Eroding20tracer}, we freeze the geometry at the second
time step from Figure~\ref{fig:Eroding20vort}, and we plot 200
trajectories that are initialized at $x=-1$ and equispaced in $y$.  The
trajectories are shown at five different times, and the final plot is a
zoom in of the lower right quadrant of the fifth time step.  Since we
use a high-order quadrature rule and a high-order time stepper, we are
able to resolve trajectories that come very close to the eroded bodies.
There are three clear regions where the tracers pass through the
geometry fastest.  Two of these regions are located between the bodies
and the solid walls at $y=\pm 1$, and the third cuts through the porous
region with the upper part of the channel formed by bodies 1, 4, 6, and
8.  Since the flow is fastest in these regions, the shearing is largest,
and this causes these openings to grow fastest which can be observed in
Figure~\ref{fig:Eroding20vort}.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270_zoom}
\caption{\label{fig:Eroding20tracer} 200 trajectories in the second
geometry from Figure~\ref{fig:Eroding20vort}. The trajectories are
initialized at $x=-1$ and are equally spaced in the $y$-direction. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot.  Since we use the Barycentric
quadrature rule and the fourth-order Runge-Kutta method, the tracers can
be very close to but not unphysically penetrate the bodies.}
\end{center}
\end{figure}

Next, we use the tracer trajectories to compute the anomalous dispersion
rates and the tortuosities in the eroding geometry.  Since these
calculations use a statistical analysis of the tracer trajectories, we
increase the number of trajectories to $N_p =
1000$~\cite{bel-sal-rin1992}.  Again, the tracers are initialized at
$x=-1$, and their $y$ coordinates are distributed evenly across the
entire channel.  To compute the tortuosity, we require the velocity at
the initialization points of the tracers.  These normalized velocities
are plotted in Figure~ref{fig:Eroding20tort}(a) for the eroded geometry
with porosity 62.9\% (see Figure~\ref{fig:Eroding20tort}(c)).  The
velocities are similar to the work of Matyka et
al.~\cite{mat-kha-koz2008} (see Figure 4(a)), except that our
cross-section does not cut through any of the bodies.  Next, in
Figure~\ref{fig:Eroding20tort}(b), we plot the local
tortuosity~\eqref{eqn:localTort} by calculating the relative length of
each trajectory as it traverses the channel from $x=-1$ to $x=1$.  The
local tortuosity ranges from 1 to 1.27, meaning that one of the tracers
travelled 27\% farther than it would have if the bodies had been absent.
The average tracer travelled 9.79\% farther than if the bodies had been
absent, meaning that the tortuosity of the eroded channel is $1.098$.
Again, comparing the local tortuosity to Matyka~\cite{mat-kha-koz2008}
(Figure 4(b)), the results are qualitatively similar. However, since our
initial cross-section does not cut through any bodies, the local
tortuosity does not have any gaps.  The discontinuities correspond to
neighboring tracers with one traveling much farther than the other.  In
Figure~\ref{fig:Eroding20tort}(c), we plot pairs of trajectories that
correspond to the ten largest jumps in the local tortuosity.
Corresponding pairs are plotted in the same color.  The discrepancy
between the trajectory lengths can be attributed to the tracers
diverging from one another as they approach opposite sides of a
stagnation point in the flow.

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =0.97\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity of an eroded geometry
with porosity 62.9\%. (a) The x-component of the velocity $u(-1, y)$
normalized by its maximum velocity of $2.98 \times 10^{-3}$. (b) The
local tortuosity $\tau(y)$ along the cross section $x = -1$.  The
spatially averaged tortuosity through the channel is $1.098$.  (c) The
trajectories resulting in the ten largest differences between
neighboring trajectories.  Neighboring trajectories have the same
color.}
\end{figure}

In Figure~\ref{fig:Eroding20tort_all}, we plot the tortuosity as a
function of the porosity from the initial porosity of $0.38$ until all
the bodies have eroded.  The tortuosity is computed with both the length
of the trajectories~\eqref{eqn:tortuosity1} (red marks) and using the
spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2}.  The red square corresponds to the
porosity in Figure~\ref{fig:Eroding20tort}(c).  The two tortuosity
calculations give similar results with any discrepancy being accounted
for by slow regions of recirculation.  As the bodies erode, this creates
wide channels where trajectories undergo only minor vertical
deflections, and this explains why the tortuosity eventually decreases
with porosity.  We computed computed lines of best fit using the
porosity-tortuosity models~\eqref{eqn:tortuosityModels}, and the power
law model gave the smallest error.  The  black
dashed line in Figure~\ref{fig:Eroding20tort_all} is the line of best
fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064}
\end{align}
with a root-mean-square error of $5.9 \times 10^{-3}$.  The initial
increase in tortuosity occurs because in the absence of erosion (left
plot in Figure~\ref{fig:Eroding20vort}), many of the trajectories, such
as those initialized between bodies 15 \& 18, only perform minor
deflections to pass through the narrow regions, albeit, very slowly.
However, as erosion starts to open the channels, the trajectories
deflect into the fast regions, such as the region above body 11, and
this increases the amount of vertical deflection, and therefore the
tortuosity.

\todo[inline]{BQ: Need to fix the references to this new figure}

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/20b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20Tranposrt} (a): The tortuosity inside of an
eroding geometry initialized with 20 bodies.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue dots)
and Lagrangian method~\eqref{eqn:tortuosity1} (red stars).  The red
square correspond to the geometry in Figure~\ref{fig:Eroding20tort}(c).
The dash line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$ as
$p=0.2064$ and the root-mean-square error is $5.9 \times 10^{-3}$. (b):
The analysis of anomalous dispersion in 20 nearly touching bodies with
respect to initial porosity (as p=37.68\%) and another six porosities
during the eroding. The dashed line correspondes to a porosity of 100\%
and the dashed-dotted lines are lines of best fit with slope 1.0644 (as
p=95.1\%), 1.0673 (as p=85.09\%), 1.0594 (as p=75.15\%), 0.9664 (as
p=65.09\%), 0.7766 (as p=55.10\%), 0.7475 (as p=45.08\%), and 0.5607 (as
p=37.68\%.}
\end{figure}

>>>>

The first two of these metrics can be defined in terms of statistical
moments of trajectories.  Depending on the application, transport at the
microscale can be modelled as pure advection~\cite{dea-qua-bir-jua2018,
leb-ded-dav-bou2007, cve-che-wen1996, puy-gou-den2019}, and advection
dispersion equation~\cite{cus-hu-den1995, dag1987, den-ica-hid2018}, or
with a random walk~\cite{saf1959, bij-blu2006, ber-sch-sil2000}.  In
this work, we assume that the trajectories are governed purely by
advection (ie.~no diffusion).  Therefore, a trajectory $\ss(t)$
initialized at $\ss_0$ is governed by
\begin{align}
  \frac{d\ss}{dt} = \uu(\ss,t), \quad \ss(0) = \ss_0,
  \label{eqn:tracers}
\end{align}
where $\uu$ is the fluid velocity.  Since there is no diffusion, all
dispersion is a result of the particle spreading induced by the porous
geometry.  Since there is no diffusion, we apply a high-order explicit
Runge-Kutta time stepping method to compute streamlines.  However, other
groups have simulated trajectories using continuous time random
walks~\cite{mon-wei1965, den-cor-sch-ber2004, leb-den-car2008,
ber-cor-den-sch2006} and a Fokker-Planck equation~\cite{ica-den2019}.

@@

The tortuosity is a first statistical moment of the streamlines.
However, to measure spread of trajectories, second-order moments are
computed. By computing the variance of the lengths of trajectories in a
porous media, non-local anomalous dispersion is
observed~\cite{kan-dea-nun-bij-blu-jua2014, cus-hu-den1995,
dea-leb-den-tar-bol-dav2013}, resulting in super-dispersion.  The
dispersion depends on the amount of time trajectories spend in low and
high velocity regimes~\cite{ber-sch2001}.  Anomalous dispersion has been
investigated for a variety of porous geometries, but not in the context
of eroding porous geometries.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport, Tracers, and Tortuosity}
\label{sec:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Erosion leads to interesting porous geometries, and we are interested in
characterizing transport in these geometries.  In our previous
work~\cite{qua-moo2018}, we examined the effect of erosion on the area
fraction, flow rate, and the total drag.  However, to characerize
transport, other quantities must be examined.  Here, we compute the
anomalous dispersion rate, the tortuosity, and the distribution of the
gap sizes.  The first two metrics can be defined in terms of streamline
governed by an advection autonomous equation~\eqref{eqn:tracers}.

@@

We calculate the anomalous dispersion rates by analyzing the streamlines
governed by equation~\eqref{eqn:tracers} for geometries that are formed
by erosion.  Given a set of trajectories, we define $\lambda_j(t)$ to be
the arclength of the trajectory
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p.
\end{align}
Then, the first and second ensemble moments are
\begin{align}
  \label{eqn:moments}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
Since we compute second-order statistics, we use 1000 streamlines so
that the statistics have converged~\cite{bel-sal-rin1992}.  Then,
particle dispersion is characterized as $\sigma_\lambda$.  At early
times, the particles have not explored much of the geometry, and we
expect a ballistic motion $\sigma_\lambda \sim t$.  However, as the
particles pass the grains, their trajectories are altered, and we expect
that $\sigma_\lambda \sim t^\alpha$, with $\alpha \in (0.5,1)$,
indicating that the flow is super-dispersive.

@@

To compute the tortuosity and the anomalous dispersion rates
(Section~\ref{sec:transport}), we require the streamlines governed by equation~\eqref{eqn:tracers}.  To reliably
characterize the transport, it is crucial that the streamlines are
computed accurately.  If a low-order time stepping method is used, or if
$\uu(\ss(t))$ is inaccurate, then a trajectory $\ss(t)$ can unphysically
enter a grain, rendering the trajectory meaningless to quantify
transport.  However, if we simply ignored trajectories that pass close
to grains, this would considerably skew the characterization of the
transport. Therefore, we use high-order quadrature and time stepping
methods.  In particular, depending on the proximity of $\ss(t)$ to
$\bd\Omega$ (Section~\ref{sec:fmm}, we apply the trapezoid rule or the
Barycentric quadrature rule.  For time stepping, we use a fourth-order
explicit Runge-Kutta method.  Since the trajectories are computed with
high-order methods in both space and time,  we are able to simulate
dynamics very close to the eroding bodies (see
Figures~\ref{fig:Eroding20tracer} and~\ref{fig:Eroding100tort}).

@@

The common characteristic of each of the experiments is near-contact
between the eroding bodies, outer walls, and streamlines. We use our
numerical methods to describe and visualize the following examples:
\begin{itemize}
  \item{\bf Single Body Close to a Wall}: We consider a single eroding
  body close to the outer wall at $y=-1$.  Using a shear flow that
  vanishes on $y=-1$, we compare the eroding body's shape to a similar
  experiment of Spagnolie and Mitchell~\cite{mit-spa2017}.

  \item{\bf 20 Bodies at a Medium Porosity}: We consider 20 eroding
  bodies with a medium initial porosity.  After computing accurate
  streamlines, the tortuosity and anomalous dispersion rates are
  computed and compared to those of an open channel.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points and a time step size of $\Delta t = 10^{-4}$.  The background
flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 2.  Snapshots of the bodies at four equispaced
times are shown in Figure~\ref{fig:Eroding20vort}.  The color is the
vorticity which is equivalent to the shear stress when restricted to the
eroding bodies.  Therefore, the rate of erosion is fastest in regions
where the magnitude of the vorticity is largest.  Initially, several of
the eroding bodies are closer to the outer wall than the $5h$ threshold
required to guarantee that the trapezoid rule achieves machine
precision.  In particular, with $h=L_2/N_\out$, the distance between
bodies 1, 6, 13, and 15 and the outer wall are $1.3h$, $2.9h$, $2.8h$,
and $1.3h$, respectively.  In addition, the distance between several
pairs of eroding bodies is too small to be accurately resolved with the
trapezoid rule at the prescribed resolution.  Particular pairs include
bodies 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18.  By using the Barycentric
quadrature rule, the interaction between these nearly-touching bodies
can be accurately resolved at the prescribed resolution, and erosion can
be simulated until all the bodies have vanished.

@@

We observe that erosion creates a network of channels from the inlet to
outlet where the velocity and vorticity, and therefore erosion rate, are
much larger when compared to other regions.  These channels can be
further visualized by considering the streamlines.  In
Figure~\ref{fig:Eroding20tracer}, we freeze the geometry at the second
time step from Figure~\ref{fig:Eroding20vort}, and we plot 200
streamlines that are initialized at $x=-1$ and equispaced in $y$.  The
streamlines are shown at five different times, and the final plot is a
zoom in of the lower right quadrant of the fifth time step.  Since we
use a high-order quadrature rule and a high-order time stepping method,
we resolve streamlines that come very close to the eroded bodies.  There
are three clear regions where the streamlines are concentrated, and this
corresponds to regions of high velocity.  Two of these regions are
located between the bodies and the solid walls at $y=\pm 1$, and the
third cuts through the porous region with the upper part of the channel
formed by bodies 1, 4, 6, and 8.  Since the flow is fastest in these
regions, the shearing is largest, and this causes these openings to grow
fastest which can be observed in Figure~\ref{fig:Eroding20vort}.

Similar to our previous work, erosion causes the space between nearly
touching bodies to quickly expand, and flat faces develop along the
region of near contact.  This qualitative behavior is present in
Figure~\ref{fig:Eroding20vort} between bodies 3 \& 4, 15 \& 16, and
others.  However, by resolving the interaction betweeen bodies that are
much closer together than in our previous work, we observe that, at
least initially, very little erosion occurs between certain pairs of
bodies.  For instance the opening between bodies 1 \& 6, 3 \& 9, and 5
\& 13 grow much slower than the opening between bodies 15 \& 16.  A
common feature of the openings that grow slowly is that they are nearly
perpendicular to the main flow direction, and this results in a small
velocity, and hence a slow erosion rate.  In fact, when plotting the
streamlines, only a small fraction of the plotted streamlines pass
between theese pairs of bodies indicating a very small velocity.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from Figure~\ref{fig:Eroding20vort}. The streamlines are
initialized at $x=-1$ and are equally spaced in the $y$-direction. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot.  Since we use the Barycentric
quadrature rule and the fourth-order Runge-Kutta method, the streamlines
can be very close to but not unphysically penetrate the bodies.}
\end{center}
\end{figure}

Next, we use the streamlines to compute the local tortuosity and the
anomalous dispersion rate of eroding geometries.  Since these
calculations are statistical, we use $N_p = 1000$
streamlines~\cite{bel-sal-rin1992}.  To compute the tortuosity, we
require the velocity at the start of the streamlines.  These normalized
velocities are plotted in Figure~\ref{fig:Eroding20tort}(a) for the
eroded geometry with porosity 62.9\% (see
Figure~\ref{fig:Eroding20tort}(c)).  The velocities are similar to the
work of Matyka et al.~\cite{mat-kha-koz2008} (see Figure 4(a)), except
that our cross-section, by construction, does not cut through any of the
bodies.  Next, in Figure~\ref{fig:Eroding20tort}(b), we plot the local
tortuosity~\eqref{eqn:localTort} by calculating the relative length of
each streamline as it traverses the channel from $x=-1$ to $x=1$.  The
local tortuosity ranges from 1 to 1.27, meaning that one of the
streamlines is 27\% longer than it would have been if the bodies were
absent.  The average streamline length is 9.79\% longer than if the
bodies had been absent, so the tortuosity of the eroded channel is
$1.098$.  Again, comparing the local tortuosity to
Matyka~\cite{mat-kha-koz2008} (Figure 4(b)), the results are
qualitatively similar. However, since our initial cross-section does not
cut through any bodies, the local tortuosity does not have any gaps.
The discontinuities in the local tortuosity happens neighboring
streamlines diverge on opposite sides of a body.  In
Figure~\ref{fig:Eroding20tort}(c), we plot pairs of streamline that
correspond to the ten largest jumps in the local tortuosity.
Corresponding pairs are plotted in the same color.

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =0.97\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity of an eroded geometry
with porosity 62.9\%. (a) The $x$-component of the velocity, $u(-1,y)$,
normalized by its maximum velocity of $2.98 \times 10^{-3}$. (b) The
local tortuosity $\tau(y)$ along the cross section $x = -1$.  The
spatially averaged tortuosity through the channel is $1.098$.  (c) The
streamlines resulting in the ten largest differences between neighboring
streamlines.  Neighboring streamlines have the same color.}
\end{figure}

In Figure~\ref{fig:Eroding20Transport}(a), we plot the tortuosity as a
function of the porosity from the initial porosity of $0.38$ until all
the bodies have eroded.  The tortuosity is computed with both the length
of the streamlines~\eqref{eqn:tortuosity1} (red marks) and using the
spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in Figure~\ref{fig:Eroding20tort}(c).
The two tortuosity formulas give similar results, and any discrepancy
can be accounted for by slow regions of recirculation and quadrature
error induced by the integrals in the tortuosity definitions.  As the
bodies erode, this creates wide channels where streamlines undergo only
minor vertical deflections, and this explains why the tortuosity
eventually decreases with porosity.  We computed lines of best fit using
the porosity-tortuosity models~\eqref{eqn:tortuosityModels}, and the
power law model gave the smallest error.  The  black dashed line in
Figure~\ref{fig:Eroding20Transport}(a) is the line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2064}
\end{align}
with a root-mean-square error of $5.9 \times 10^{-3}$.  The initial
increase in tortuosity occurs because in the absence of erosion (left
plot in Figure~\ref{fig:Eroding20vort}), many of the streamlines, such
as those initialized between bodies 15 \& 18, only perform minor
deflections to pass through the narrow regions, albeit, very slowly.
However, as erosion starts to open the channels, the streamlines deflect
into the fast regions, such as the region above body 11, and this
increases the amount of vertical deflection, and therefore the
tortuosity.

\todo[inline]{Nothing about the anomalous dispersion rate here}
At high porosities, the dispersion rate is close to 1, and this can be
caused by one of two reasons.  1) There is no measurable anomalous
dispersion; or, 2) The streamlines are too short to see the anomalous
dispersion.

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/20b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20Transport} (a): The tortuosity inside of an
eroding geometry initialized with 20 bodies.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue dots)
and Lagrangian method~\eqref{eqn:tortuosity1} (red stars).  The red
square correspond to the geometry in Figure~\ref{fig:Eroding20tort}(c).
The dash line is the line of best fit $\widehat{T}(\phi)=\phi^{-p}$ with
$p=0.2064$, and the corresponding root-mean-square error is $5.9 \times
10^{-3}$. (b): The temporal evolution of $\sigma_\lambda$ as a function
of time at seven different porosities.  The dashed, which has slope one,
corresponds to $\sigma_\lambda$ if no bodies were present.  At early
times, the particles undergo a ballistic motion, but once they have
traversed a few grains, the dispersion becomes super-dispersive.  The
dashed-dotted lines are lines of best fit with slopes $1.0644$
($\phi=95.1\%$), $1.0673$ ($\phi=85.09\%$), $1.0594$ ($\phi=75.15\%$),
$0.9664$ ($\phi=65.09\%$), $0.7766$ ($\phi=55.10\%$), $0.7475$
($\phi=45.08\%$), and $0.5607$ ($\phi=37.68\%$).}
\end{figure}
